---
title: "CSN_lab2"
output: html_notebook
---

# Analysis of nodes degree distribution 

```{r}
library(ggplot2)
library(xtable)
require("stats4")
require("VGAM")
library(reshape2)
library(gridExtra) 
```


# IN-DEGREE DISTRIBUTION

### Collect the node degree of every language

```{r}
source = read.table("list_in.txt", header = TRUE, 
         as.is = c("language","file"))

degree_distrib_ls = c()

for (x in 1:nrow(source)) {
    degree_distrib_ls[x] = read.table(source$file[x])
}

```


## Summary Table

```{r}
summary_tab <- data.frame()
i = 1
for (l in 1:length(degree_distrib_ls)) {
  summary_tab[i,1] <- source$language[i]
  summary_tab[i,2] <- length(unlist(degree_distrib_ls[[l]]))
  summary_tab[i,3] <- max(unlist(degree_distrib_ls[[l]]))
  summary_tab[i,4] <- round(sum(unlist(degree_distrib_ls[[l]]))/length(unlist(degree_distrib_ls[[l]])),2)
  summary_tab[i,5] <- round(length(unlist(degree_distrib_ls[[l]]))/sum(unlist(degree_distrib_ls[[l]])),2)
  i = i + 1
}

colnames(summary_tab) <- c("lang","N","max k","M/N","N/M")
#print(xtable(summary_tab))
summary_tab
```

### summary of summary

```{r}
print(xtable(summary(summary_tab)))
```



# Fitting Distributions

## Opt parameters with MLE
Get a vector of optimal parameters for each language, and the value of 2mloglik

```{r}

get_mle_params <- function(x) {
    # global variables
    M <- sum(x)
    N <- length(x)
    M_1 <- sum(log(x))
    H <- function(kmax,gamma) {
      s <- 0
      for (i in 1:kmax) {
        s <- s + (1/(i^gamma))
        }
      return(s)
      }
    
    ## Poisson
    minus_log_like_poisson <- function(lambda) {
      C <- 0;
      for(i in 1:N) {
        for(j in 2:x[i]) {
          C<-C+log(j)
        }
      }
      return(- M * log(lambda) + N*(lambda + log(1-(exp(-lambda)))) + C)
    }
    
    
    ## Geometric
    minus_log_like_geom <- function(q) {
      - (M-N)*log(1-q) - N*log(q)
    }
    
    ## zeta
    minus_log_like_zeta <- function(gamma) {
      N * log(zeta(gamma)) + gamma * M_1
    }

    ## Trunc Zeta
    minus_log_like_trunc_zeta <- function(kmax,gamma) {
      gamma*M_1 + N*log(H(kmax,gamma))
    }

    # ml estimation
    
    mle_geom <- mle(minus_log_like_geom,start = list(q = N/M),method = "L-BFGS-B",
                    lower = c(0.00001),upper = c(0.99999))
    
    mle_zeta <- mle(minus_log_like_zeta,start = list(gamma = 5),method = "L-BFGS-B",
                    lower = c(1.0000001))
    
    mle_poisson <- mle(minus_log_like_poisson,start = list(lambda = M/N),method = "L-BFGS-B",
                       lower = c(0.0000001))
    
    mle_trunc_zeta <- mle(minus_log_like_trunc_zeta,start = list(gamma = 3, kmax=max(x)),method = "L-BFGS-B",
                       lower = list(kmax = 1), upper = list(kmax = N))
    
    
    mle_ls <- c(mle_poisson,mle_geom,mle_zeta,mle_trunc_zeta)
    
    
    
    # collect the best par for each distribution
    best_pars <- c()
    i <- 1
    for (n in 1:length(mle_ls)) {
      if (n < length(mle_ls)) {
          best_pars[i] <- attributes(summary(mle_ls[[n]]))$coef[1]
          i <- i + 1
          }
      else {
        best_pars[i] <- attributes(summary(mle_ls[[n]]))$coef[2]
        best_pars[i+1] <- attributes(summary(mle_ls[[n]]))$coef[1]
      }
      
    }
    
    
    # collect -2logLike 
    m2loglik_ls <- c()
    i <- 1
    for (n in mle_ls) {
      m2loglik_ls[i] <- m2logL <- attributes(summary(n))$m2logL
      i <- i + 1
    }

    
    return(list(best_pars,m2loglik_ls))
    
}

```


### Optimal Parameters table
```{r}
params_table <- data.frame()

for (n in 1:length(degree_distrib_ls)) {
    x <- unlist(degree_distrib_ls[[n]])
    new_pars <- get_mle_params(x)
    params_table <- rbind(params_table,new_pars[[1]])
}

params_table$lang <- source$language
colnames(params_table) <- c("lambda","q","gamma_1","gamma_2","kmax","lang")

params_table
#print(xtable(params_table))
```


# Model Selection

## AIC Table

```{r}
get_AIC <- function(m2logL,K,N) {
   m2logL + 2*K*N/(N-K-1) # AIC with a correction for sample size
}


aic_table <- data.frame()

for (n in 1:length(degree_distrib_ls)) {
    x <- unlist(degree_distrib_ls[[n]])
    new_pars <- get_mle_params(x)
    
    N <- length(x)
    M <- sum(x)
    M_1 <- sum(log(x))
    minus_log_likelihood_zeta2 <- 2*M_1 + N * log((pi^2)/6)
    
    aic_poiss <- get_AIC(new_pars[[2]][1],1,N)
    aic_geom <- get_AIC(new_pars[[2]][2],1,N)
    aic_zeta <- get_AIC(new_pars[[2]][3],1,N)
    aic_trunc_zeta <- get_AIC(new_pars[[2]][4],2,N)
    aic_zeta_2 <- get_AIC(2*minus_log_likelihood_zeta2,0,N)
    
    aic_ls <- c(aic_poiss,aic_geom,aic_zeta,aic_trunc_zeta,aic_zeta_2)
    aic_table <- rbind(aic_table,aic_ls)
    aic_table$best[n] <- which.min(aic_table[n,1:5])
}

aic_table$lang <- source$language
colnames(aic_table) <- c("Poisson","Geometric","Zeta","Trunc_Zeta","Zeta_2","best","lang")
aic_table
#print(xtable(aic_table))

```

### AIC diff table

```{r}
aic_diff <- data.frame()
for (i in 1:nrow(aic_table)) {
  aic_diff[i,1] <- aic_table[i,1]-min(aic_table[i,1:5])
  aic_diff[i,2] <- aic_table[i,2]-min(aic_table[i,1:5])
  aic_diff[i,3] <- aic_table[i,3]-min(aic_table[i,1:5])
  aic_diff[i,4] <- aic_table[i,4]-min(aic_table[i,1:5])
  aic_diff[i,5] <- aic_table[i,5]-min(aic_table[i,1:5])
  i = i + 1
}

aic_diff$lang <- source$language
colnames(aic_diff) <- c("Poisson diff AIC","Geometric diff AIC","Zeta diff AIC","Zeta Trunc diff AIC","Zeta 2 diff AIC","lang")
aic_diff
#print(xtable(aic_diff))
aic_diff
```


### Null and Target models

```{r}
mean_aic_comparison <- data.frame()
i = 1
for (i in 1:nrow(aic_table)) {
  mean_aic_comparison[i,1] <- aic_table[i,1]
  mean_aic_comparison[i,2] <- aic_table[i,2]
  mean_aic_comparison[i,3] <- mean(c(aic_table[i,3],aic_table[i,4],aic_table[i,5]))
  i = i + 1
}

mean_aic_comparison$lang <- source$language
colnames(mean_aic_comparison) <- c("Poisson AIC","Geometric AIC","Mean Zeta AIC","lang")
mean_aic_comparison
#print(xtable(mean_aic_comparison))

```


### plot AIC and N

```{r}
library(ggpubr)
setEPS()
postscript("N_AIC.eps",width = 6, height = 3)

a <- ggplot(data=summary_tab, aes(x=reorder(lang, -N), y=N)) +
  geom_bar(stat="identity", fill = "magenta") +
  geom_text(aes(label=N), vjust=1.6,size=2.5) +
  theme_minimal() +
  labs(y= "number of nodes",x= "language") +
  theme(axis.title=element_text(size=10),axis.text.x=element_text(angle=45,hjust=1)) +
  coord_flip()

melt_measures <- melt(mean_aic_comparison,id = "lang")
b <- ggplot(melt_measures, aes( x=reorder(lang, -value),y = value, colour=variable, group=variable )) + 
  geom_line() +
  labs(x="language",y="AIC") +
  coord_flip() +
  theme(axis.text.x=element_text(angle=45,hjust=1),
        legend.key.size = unit(0.8, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title = element_text(size=12), #change legend title font size
        legend.text = element_text(size=10), #change legend text font size
        legend.position = c(0.8,0.8),
        axis.title=element_text(size=10))

ggarrange(a,b,nrow=1,ncol=2)

dev.off()


```


## Visualization

### Define distributions

```{r}
# real distribution 
real_dist <- function(x){
  degree_spectrum <- table(x)
  N <- length(x)
  degree_dist <- data.frame(degree_spectrum/N)
  return(degree_dist)
}

# ZETA
estimated_dist_zeta <- function(xvals,gamma){
  return(1/(zeta(gamma)*xvals^gamma)); # definition of the zeta-function
}

# TRUNC ZETA
H <- function(kmax,gamma) {
  s <- 0
  for (i in 1:kmax) {
    s <- s + (1/(i^gamma))
  }
  return(s)
}

zeta_trunc_dist <- function(xvals,gamma,kmax){
  return(1/(H(kmax,gamma)*xvals^gamma)); # definition of the zeta-trunc-function
}

poisson_dist <- function(xvals,q) {
  return((1 - q)^(xvals - 1)*q)
}

```


### Plots

```{r}
plot_distrib_zeta <- function(x,lang) {
  xvals <- sort(as.integer(unique(x)))
  y_zeta <- estimated_dist_zeta(gamma = params_table$gamma_1[params_table$lang == lang],xvals = xvals)
  y_real <- real_dist(x)
  
  dist_zeta <- data.frame("y_pred"= y_zeta, "y_real" = y_real$Freq, "degree" = as.integer(y_real$x))
  
  a <- ggplot(data=dist_zeta) +
    geom_bar(aes(x=degree, y=y_pred),stat="identity", fill = "magenta") +
    geom_bar(aes(x=degree, y=y_real),stat="identity", fill = "green") +
    geom_point(aes(x=degree, y=y_pred, colour = "red")) +
    geom_point(aes(x=degree, y=y_real, colour = "green")) +
    theme_minimal() +
    labs(x="k",y="p(k)", colour = "Legend \n",title=paste( lang, "- Zeta")) +
    theme(legend.position = c(0.2,0.2)) +
    scale_color_manual(labels = c("Real values", "Estimated values"), values = c("green", "magenta")) +
    scale_x_log10(labels=scales::trans_format('log10',scales::math_format(10^.x))) +
    scale_y_log10(labels=scales::trans_format('log10',scales::math_format(10^.x)))
  return(a)
}


plot_distrib_zeta_trunc <- function(x,lang) {
  xvals <- sort(as.integer(unique(x)))
  y_zeta_trunc <- zeta_trunc_dist(xvals,
                                  gamma = params_table$gamma_2[params_table$lang == lang],
                                  kmax = params_table$kmax[params_table$lang == lang])
  y_real <- real_dist(x)
  
  dist_zeta_trunc <- data.frame("y_pred"= y_zeta_trunc, "y_real" = y_real$Freq, "degree" = as.integer(y_real$x))
  
  a <- ggplot(data=dist_zeta_trunc) +
    geom_bar(aes(x=degree, y=y_pred),stat="identity", fill = "magenta") +
    geom_bar(aes(x=degree, y=y_real),stat="identity", fill = "green") +
    geom_point(aes(x=degree, y=y_pred, colour = "red")) +
    geom_point(aes(x=degree, y=y_real, colour = "green")) +
    theme_minimal() +
    labs(x="k",y="p(k)", colour = "Legend \n",title=paste( lang, "- Truncated Zeta")) +
    theme(legend.position = c(0.2,0.2)) +
    scale_color_manual(labels = c("Real values", "Estimated values"), values = c("green", "magenta")) +
    scale_x_log10(labels=scales::trans_format('log10',scales::math_format(10^.x))) +
    scale_y_log10(labels=scales::trans_format('log10',scales::math_format(10^.x)))
  return(a)
}

plot_distrib_poisson <- function(x,lang) {
  xvals <- sort(as.integer(unique(x)))
  y_poisson <- poisson_dist(q = params_table$q[params_table$lang == lang],xvals = xvals)
  y_real <- real_dist(x)
  
  dist_poisson <- data.frame("y_pred"= y_poisson, "y_real" = y_real$Freq, "degree" = as.integer(y_real$x))
  
  a <- ggplot(data=dist_poisson) +
    geom_bar(aes(x=degree, y=y_pred),stat="identity", fill = "magenta") +
    geom_bar(aes(x=degree, y=y_real),stat="identity", fill = "green") +
    geom_point(aes(x=degree, y=y_pred, colour = "red")) +
    geom_point(aes(x=degree, y=y_real, colour = "green")) +
    theme_minimal() +
    labs(x="k",y="p(k)", colour = "Legend \n",title=paste( lang, "- Poisson")) +
    theme(legend.position = c(0.2,0.2)) +
    scale_color_manual(labels = c("Real values", "Estimated values"), values = c("green", "magenta")) +
    scale_x_log10(labels=scales::trans_format('log10',scales::math_format(10^.x))) +
    scale_y_log10(labels=scales::trans_format('log10',scales::math_format(10^.x)))
  return(a)
}



setEPS()
postscript("try1.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[1]])
plot_distrib_zeta_trunc(x,"Arabic")
dev.off()

setEPS()
postscript("try2.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[2]])
plot_distrib_zeta_trunc(x,"Basque")
dev.off()

setEPS()
postscript("try3.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[3]])
plot_distrib_zeta_trunc(x,"Catalan")
dev.off()

setEPS()
postscript("try4.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[4]])
plot_distrib_zeta_trunc(x,"Chinese")
dev.off()

setEPS()
postscript("try5.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[5]])
plot_distrib_zeta_trunc(x,"Czech")
dev.off()

setEPS()
postscript("try6.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[6]])
plot_distrib_zeta_trunc(x,"English")
dev.off()

setEPS()
postscript("try7.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[7]])
plot_distrib_zeta_trunc(x,"Greek")
dev.off()

setEPS()
postscript("try8.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[8]])
plot_distrib_zeta(x,"Hungarian")
dev.off()

setEPS()
postscript("try9.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[9]])
plot_distrib_zeta_trunc(x,"Italian")
dev.off()

setEPS()
postscript("try10.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[10]])
plot_distrib_zeta(x,"Turkish")
dev.off()

# poisson
setEPS()
postscript("try11.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[5]])
plot_distrib_poisson(x,"Czech")
dev.off()


```




# Alternative Models

## Sonja model
A right truncated Zeta for k < threshold, and 1/N otherwise

```{r}
sonja_dist <- function(x,gamma,kmax){
  N <- length(x)
  xvals <- sort(as.integer(unique(x)))
  
  p_list <- c()
  i=1 
  for (k in xvals) {
    if (k <= 100) {
      p_list[i] <- (1/(H(kmax,gamma)*k^gamma)) # definition of the zeta-trunc-function
    }
    else {
      p_list[i] <- (1/N)
    }
    i = i + 1
  }
  return(p_list)
}

plot_distrib_sonja <- function(x,lang) {
  xvals <- sort(as.integer(unique(x)))
  y_pred <- sonja_dist(x,
                       gamma = params_table$gamma_2[params_table$lang == lang],
                       kmax = params_table$kmax[params_table$lang == lang])
  y_real <- real_dist(x)
  
  dist_sonja <- data.frame("y_pred"= y_pred, "y_real" = y_real$Freq, "degree" = as.integer(y_real$x))
  
  a <- ggplot(data=dist_sonja) +
    geom_bar(aes(x=degree, y=y_pred),stat="identity", fill = "magenta") +
    geom_bar(aes(x=degree, y=y_real),stat="identity", fill = "green") +
    geom_point(aes(x=degree, y=y_pred, colour = "red")) +
    geom_point(aes(x=degree, y=y_real, colour = "green")) +
    theme_minimal() +
    labs(x="k",y="p(k)", colour = "Legend \n",title=paste( lang, "- Truncated Zeta")) +
    theme(legend.position = c(0.2,0.2)) +
    scale_color_manual(labels = c("Real values", "Estimated values"), values = c("green", "magenta")) +
    scale_x_log10(labels=scales::trans_format('log10',scales::math_format(10^.x))) +
    scale_y_log10(labels=scales::trans_format('log10',scales::math_format(10^.x)))
  return(a)
}

x <- unlist(degree_distrib_ls[[7]])
plot_distrib_sonja(x,"Italian")
```

## Trying to optimize parameters

```{r}
get_mle_pars_sonja <- function(x) {
      M <- sum(x)
      N <-  length(x)
      M_1 <- sum(log(x))
        
      H <- function(kmax,gamma) {
      s <- 0
      for (i in 1:kmax) {
        s <- s + (1/(i^gamma))
        }
      return(s)
      }
      
      # Loglikelihood
      minus_log_like_sonja <- function(gamma,t) {
          xvals <- sort(as.integer(unique(x)))
        
          for (k in xvals) {
            if (k <= t) {
              return(gamma*M_1 + N*log(H(kmax,gamma))) # definition of the zeta-trunc-function
            }
            else {
              return(N(log(1/N)))
            }
            i = i + 1
          }
      }
      
      mle_alt <- mle(minus_log_like_sonja, start = list(gamma = 2, kmax = max(x), t = 90),method = "L-BFGS-B",
                     lower = list(kmax = 1), upper = list(kmax = N))
      
      return(list(attributes(summary(mle_alt))$coef[1],
                  attributes(summary(mle_alt))$coef[2],
                  attributes(summary(mle_alt))$m2logL))
}

x <- unlist(degree_distrib_ls[[3]])
sonja_pars <- get_mle_pars_sonja(x)
```






## Altman's function

```{r}
get_mle_pars_altman <- function(x) {
      M <- sum(x)
      M_1 <- sum(log(x))
          
      # Loglikelihood
      S <- function(gamma,delta) {
            s <- 0
            for (i in 1:N) {
              s <- s + (i^(-gamma)*exp(-delta*i))
              }
            return(s) 
      }
      
      minus_log_like_alt <- function(gamma,delta) {
            gamma*M_1 + delta*M + log(S(gamma,delta))
      }
      
      mle_alt <- mle(minus_log_like_alt, start = list(gamma = 2, delta = -3),method = "L-BFGS-B")
      
      return(list(attributes(summary(mle_alt))$coef[1],
                  attributes(summary(mle_alt))$coef[2],
                  attributes(summary(mle_alt))$m2logL))
}

x <- unlist(degree_distrib_ls[[1]])
alt_pars <- get_mle_pars_altman(x)
get_AIC(alt_pars[[3]][1],2,N)
```


```{r}
aic_table <- data.frame()

for (n in 1:length(degree_distrib_ls)) {
    x <- unlist(degree_distrib_ls[[n]])
    new_pars <- get_mle_params(x)
    alt_pars <- get_mle_pars_altman(x)
    
    N <- length(x)
    M <- sum(x)
    M_1 <- sum(log(x))
    minus_log_likelihood_zeta2 <- 2*M_1 + N * log((pi^2)/6)
    
    aic_poiss <- get_AIC(new_pars[[2]][1],1,N)
    aic_geom <- get_AIC(new_pars[[2]][2],1,N)
    aic_zeta <- get_AIC(new_pars[[2]][3],1,N)
    aic_trunc_zeta <- get_AIC(new_pars[[2]][4],2,N)
    aic_zeta_2 <- get_AIC(2*minus_log_likelihood_zeta2,0,N)
    aic_alt <- get_AIC(alt_pars[[3]][1],2,N)
    
    aic_ls <- c(aic_poiss,aic_geom,aic_zeta,aic_trunc_zeta,aic_zeta_2,aic_alt)
    aic_table <- rbind(aic_table,aic_ls)
    aic_table$best[n] <- which.min(aic_table[n,1:6])
}

aic_table$lang <- source$language
colnames(aic_table) <- c("Poisson","Geometric","Zeta","Trunc_Zeta","Zeta_2","Altman","best","lang")
aic_table
```



### compare with real distributions
```{r}
degree_sequence = read.table("./samples_from_discrete_distributions/data/sample_of_geometric_with_parameter_0.2.txt",
                             header = FALSE)
```









