---
title: "CSN_lab2"
output: html_notebook
---

# Analysis of nodes degree distribution 

After finding the parameters and the best model, plot the data and a theoretical curve to see if there is sufficient agreement 

("I give x and y, use a double log scale for viz")

```{r}
library(ggplot2)
```


Degree sequence of the English network

```{r}
degree_sequence = read.table("./data/English_degree_sequence.txt",
                  header = FALSE)
nrow(degree_sequence)   # number of nodes 
```
```{r}
sum(degree_sequence)    # sum of total degrees
sum(degree_sequence)/dim(degree_sequence)[1]    # mean degree
spectrum <- table(degree_sequence)
```


```{r}
barplot(spectrum, main = "English",
        xlab = "degree", ylab = "number of vertices")
```

```{r}
barplot(spectrum, main = "English",
        xlab = "degree", ylab = "number of vertices", log = "xy")
```


```{r}
source("summary_table.R")
```


# IN-DEGREE DISTRIBUTION

Table 1: Summary of the properties of the degree sequences. N is the number of nodes, M/N is mean degree where M is the sum of degrees.

```{r}
source = read.table("list_in.txt", header = TRUE, 
         as.is = c("language","file"))

for (x in 1:nrow(source)) {
    write_summary(source$language[x], source$file[x])
}
```








