---
title: "CSN_lab2"
output: html_notebook
---

# Analysis of nodes degree distribution 

After finding the parameters and the best model, plot the data and a theoretical curve to see if there is sufficient agreement 

("I give x and y, use a double log scale for viz")

```{r}
library(ggplot2)
library(xtable)
require("stats4") # for MLE
require("VGAM")
library(reshape2)
library(gridExtra) 
```


Degree sequence of the English network

```{r}


#nrow(degree_sequence)   # number of nodes 
#sum(degree_sequence)    # sum of total degrees
#sum(degree_sequence)/dim(degree_sequence)[1]    # mean degree
#spectrum <- table(degree_sequence)
```

```{r}
#barplot(spectrum, main = "English",
#        xlab = "degree", ylab = "number of vertices")
```

```{r}
#barplot(spectrum, main = "English",
#        xlab = "degree", ylab = "number of vertices", log = "xy")
```


# IN-DEGREE DISTRIBUTION

### Collect the node degree of every language

```{r}
source = read.table("list_in.txt", header = TRUE, 
         as.is = c("language","file"))

degree_distrib_ls = c()

for (x in 1:nrow(source)) {
    degree_distrib_ls[x] = read.table(source$file[x])
}


```


## Summary Table

```{r}
summary_tab <- data.frame()
i = 1
for (l in 1:length(degree_distrib_ls)) {
  summary_tab[i,1] <- source$language[i]
  summary_tab[i,2] <- length(unlist(degree_distrib_ls[[l]]))
  summary_tab[i,3] <- max(unlist(degree_distrib_ls[[l]]))
  summary_tab[i,4] <- round(sum(unlist(degree_distrib_ls[[l]]))/length(unlist(degree_distrib_ls[[l]])),2)
  summary_tab[i,5] <- round(length(unlist(degree_distrib_ls[[l]]))/sum(unlist(degree_distrib_ls[[l]])),2)
  i = i + 1
}

colnames(summary_tab) <- c("lang","N","max k","M/N","N/M")
#print(xtable(summary_tab))
summary_tab
```

### summary of summary

```{r}
print(xtable(summary(summary_tab)))
```



## Fitting Distributions

### Find starting parameter
# For gammma_1 and gamma_2, we can fit a lin reg of log(f(k)) over log(k), we measure probability thorugh sequence 


### Opt parameters with MLE

```{r}

get_mle_params <- function(x) {
    # global variables
    M <- sum(x)
    N <- length(x)
    M_1 <- sum(log(x))
    H <- function(kmax,gamma) {
      s <- 0
      for (i in 1:kmax) {
        s <- s + (1/(i^gamma))
        }
      return(s)
      }
    
    ## Poisson
    minus_log_like_poisson <- function(lambda) {
      C <- 0;
      for(i in 1:N) {
        for(j in 2:x[i]) {
          C<-C+log(j)
        }
      }
      return(- M * log(lambda) + N*(lambda + log(1-(exp(-lambda)))) + C)
    }
    
    
    ## Geometric
    minus_log_like_geom <- function(q) {
      - (M-N)*log(1-q) - N*log(q)
    }
    
    ## zeta
    minus_log_like_zeta <- function(gamma) {
      N * log(zeta(gamma)) + gamma * M_1
    }
    
    ## Trunc Zeta
    minus_log_like_trunc_zeta <- function(kmax,gamma) {
      gamma*M_1 + N*log(H(kmax,gamma))
    }

    # ml estimation
    
    mle_geom <- mle(minus_log_like_geom,start = list(q = N/M),method = "L-BFGS-B",
                    lower = c(0.00001),upper = c(0.99999))
    
    mle_zeta <- mle(minus_log_like_zeta,start = list(gamma = 2),method = "L-BFGS-B",
                    lower = c(1.0000001))
    
    mle_poisson <- mle(minus_log_like_poisson,start = list(lambda = M/N),method = "L-BFGS-B",
                       lower = c(0.0000001))
    
    mle_trunc_zeta <- mle(minus_log_like_trunc_zeta,start = list(gamma = 3, kmax=max(x)),method = "L-BFGS-B",
                       lower = list(kmax = 1), upper = list(kmax = N))
    
    
    mle_ls <- c(mle_poisson,mle_geom,mle_zeta,mle_trunc_zeta)
    
    
    
    # collect the best par for each distribution
    best_pars <- c()
    i <- 1
    for (n in 1:length(mle_ls)) {
      if (n < length(mle_ls)) {
          best_pars[i] <- attributes(summary(mle_ls[[n]]))$coef[1]
          i <- i + 1
          }
      else {
        best_pars[i] <- attributes(summary(mle_ls[[n]]))$coef[2]
        best_pars[i+1] <- attributes(summary(mle_ls[[n]]))$coef[1]
      }
      
    }
    
    
    # collect -2logLike 
    m2loglik_ls <- c()
    i <- 1
    for (n in mle_ls) {
      m2loglik_ls[i] <- m2logL <- attributes(summary(n))$m2logL
      i <- i + 1
    }

    
    return(list(best_pars,m2loglik_ls))
    
}

```


### Lang-Par table
```{r}
params_table <- data.frame()

for (n in 1:length(degree_distrib_ls)) {
    x <- unlist(degree_distrib_ls[[n]])
    new_pars <- get_mle_params(x)
    params_table <- rbind(params_table,new_pars[[1]])
}

params_table$lang <- source$language
colnames(params_table) <- c("lambda","q","gamma_1","gamma_2","kmax","lang")

params_table
#print(xtable(params_table))
```


# Model Selection

## Compute AICs

```{r}
get_AIC <- function(m2logL,K,N) {
   m2logL + 2*K*N/(N-K-1) # AIC with a correction for sample size
}


aic_table <- data.frame()

for (n in 1:length(degree_distrib_ls)) {
    x <- unlist(degree_distrib_ls[[n]])
    new_pars <- get_mle_params(x)
    
    N <- length(x)
    M <- sum(x)
    M_1 <- sum(log(x))
    minus_log_likelihood_zeta2 <- 2*M_1 + N * log((pi^2)/6)
    
    aic_poiss <- get_AIC(new_pars[[2]][1],1,N)
    aic_geom <- get_AIC(new_pars[[2]][2],1,N)
    aic_zeta <- get_AIC(new_pars[[2]][3],1,N)
    aic_trunc_zeta <- get_AIC(new_pars[[2]][4],2,N)
    aic_zeta_2 <- get_AIC(2*minus_log_likelihood_zeta2,0,N)
    
    aic_ls <- c(aic_poiss,aic_geom,aic_zeta,aic_trunc_zeta,aic_zeta_2)
    aic_table <- rbind(aic_table,aic_ls)
    aic_table$best[n] <- which.min(aic_table[n,1:5])
}

aic_table$lang <- source$language
colnames(aic_table) <- c("Poisson","Geometric","Zeta","Trunc_Zeta","Zeta_2","best","lang")
aic_table
#print(xtable(aic_table))

```

### AIC diff table

```{r}
aic_diff <- data.frame()
for (i in 1:nrow(aic_table)) {
  aic_diff[i,1] <- aic_table[i,1]-min(aic_table[i,1:5])
  aic_diff[i,2] <- aic_table[i,2]-min(aic_table[i,1:5])
  aic_diff[i,3] <- aic_table[i,3]-min(aic_table[i,1:5])
  aic_diff[i,4] <- aic_table[i,4]-min(aic_table[i,1:5])
  aic_diff[i,5] <- aic_table[i,5]-min(aic_table[i,1:5])
  i = i + 1
}

aic_diff$lang <- source$language
colnames(aic_diff) <- c("Poisson diff AIC","Geometric diff AIC","Zeta diff AIC","Zeta Trunc diff AIC","Zeta 2 diff AIC","lang")
aic_diff
#print(xtable(aic_diff))
aic_diff
```

### Null and Target models 

```{r}
mean_aic_comparison <- data.frame()
i = 1
for (i in 1:nrow(aic_table)) {
  mean_aic_comparison[i,1] <- aic_table[i,1]
  mean_aic_comparison[i,2] <- aic_table[i,2]
  mean_aic_comparison[i,3] <- mean(c(aic_table[i,3],aic_table[i,4],aic_table[i,5]))
  i = i + 1
}

mean_aic_comparison$lang <- source$language
colnames(mean_aic_comparison) <- c("Poisson AIC","Geometric AIC","Mean Zeta AIC","lang")
mean_aic_comparison
#print(xtable(mean_aic_comparison))

```

### AIC diff plot

```{r}
melt_measures <- melt(mean_aic_comparison, id.vars="lang")
setEPS()
postscript("mean_aic.eps",width = 7, height = 5)
ggplot(melt_measures, aes( x=reorder(lang, -value),y = value, colour=variable, group=variable )) + 
  geom_line() +
  labs(x="language",y="AIC") +
  coord_flip() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  theme(legend.position = c(0.8,0.8))
dev.off()
```

### plot AIC and N

```{r}
library(ggpubr)
setEPS()
postscript("N_AIC.eps",width = 6, height = 3)

a <- ggplot(data=summary_tab, aes(x=reorder(lang, -N), y=N)) +
  geom_bar(stat="identity", fill = "magenta") +
  geom_text(aes(label=N), vjust=1.6,size=2.5) +
  theme_minimal() +
  labs(y= "number of nodes",x= "language") +
  theme(axis.title=element_text(size=10),axis.text.x=element_text(angle=45,hjust=1)) +
  coord_flip()

b <- ggplot(melt_measures, aes( x=reorder(lang, -value),y = value, colour=variable, group=variable )) + 
  geom_line() +
  labs(x="language",y="AIC") +
  coord_flip() +
  theme(axis.text.x=element_text(angle=45,hjust=1),
        legend.key.size = unit(0.8, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title = element_text(size=12), #change legend title font size
        legend.text = element_text(size=10), #change legend text font size
        legend.position = c(0.8,0.8),
        axis.title=element_text(size=10))

ggarrange(a,b,nrow=1,ncol=2)

dev.off()


```


## Visualization

### Define distributions

```{r}
# real distribution 
real_dist <- function(x){
  degree_spectrum <- table(x)
  N <- length(x)
  degree_dist <- data.frame(degree_spectrum/N)
  return(degree_dist)
}

# ZETA
estimated_dist_zeta <- function(xvals,gamma){
  return(1/(zeta(gamma)*xvals^gamma)); # definition of the zeta-function
}

# TRUNC ZETA
H <- function(kmax,gamma) {
  s <- 0
  for (i in 1:kmax) {
    s <- s + (1/(i^gamma))
  }
  return(s)
}

zeta_trunc_dist <- function(xvals,gamma,kmax){
  return(1/(H(kmax,gamma)*xvals^gamma)); # definition of the zeta-trunc-function
}
```


### Plots

```{r}
plot_distrib_zeta <- function(x,lang) {
  xvals <- sort(as.integer(unique(x)))
  y_zeta <- estimated_dist_zeta(gamma = params_table$gamma_1[params_table$lang == lang],xvals = xvals)
  y_real <- real_dist(x)
  
  dist_zeta <- data.frame("y_pred"= y_zeta, "y_real" = y_real$Freq, "degree" = as.integer(y_real$x))
  
  a <- ggplot(data=dist_zeta) +
    geom_bar(aes(x=degree, y=y_pred),stat="identity", fill = "magenta") +
    geom_bar(aes(x=degree, y=y_real),stat="identity", fill = "green") +
    geom_point(aes(x=degree, y=y_pred), colour = "red") +
    geom_point(aes(x=degree, y=y_real), colour = "green") +
    theme_minimal() +
    labs(y= "log(p(k))",x="log(k)",title=paste( lang, "- Zeta")) +
    scale_x_log10() +
    scale_y_log10()
  return(a)
}


plot_distrib_zeta_trunc <- function(x,lang) {
  xvals <- sort(as.integer(unique(x)))
  y_zeta_trunc <- zeta_trunc_dist(xvals,
                                  gamma = params_table$gamma_2[params_table$lang == lang],
                                  kmax = params_table$kmax[params_table$lang == lang])
  y_real <- real_dist(x)
  
  dist_zeta_trunc <- data.frame("y_pred"= y_zeta_trunc, "y_real" = y_real$Freq, "degree" = as.integer(y_real$x))
  
  a <- ggplot(data=dist_zeta_trunc) +
    geom_bar(aes(x=degree, y=y_pred),stat="identity", fill = "magenta") +
    geom_bar(aes(x=degree, y=y_real),stat="identity", fill = "green") +
    geom_point(aes(x=degree, y=y_pred), colour = "red") +
    geom_point(aes(x=degree, y=y_real), colour = "green") +
    theme_minimal() +
    labs(y= "log(p(k))",x="log(k)",title=paste( lang, "- Truncated Zeta")) +
    scale_x_log10() +
    scale_y_log10()
  return(a)
}



plotlist <- list(a,b)
setEPS()
postscript("try1.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[1]])
a <- plot_distrib_zeta_trunc(x,"Arabic")
dev.off()

plotlist <- list(c,d)
setEPS()
postscript("try2.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[2]])
b <- plot_distrib_zeta_trunc(x,"Basque")
dev.off()

plotlist <- list(e,f)
setEPS()
postscript("try3.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[3]])
c <- plot_distrib_zeta_trunc(x,"Catalan")
dev.off()

plotlist <- list(g,h)
setEPS()
postscript("try4.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[4]])
d <- plot_distrib_zeta_trunc(x,"Chinese")

dev.off()

plotlist <- list(i,l)
setEPS()
postscript("try5.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[5]])
e <- plot_distrib_zeta_trunc(x,"Czech")
dev.off()

setEPS()
postscript("try6.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[6]])
f <- plot_distrib_zeta_trunc(x,"English")
dev.off()

setEPS()
postscript("try7.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[7]])
g <- plot_distrib_zeta_trunc(x,"Greek")
dev.off()

setEPS()
postscript("try8.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[8]])
h <- plot_distrib_zeta(x,"Hungarian")
dev.off()

setEPS()
postscript("try9.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[9]])
i <- plot_distrib_zeta_trunc(x,"Italian")
dev.off()

setEPS()
postscript("try10.eps",width = 6, height = 3)
x <- unlist(degree_distrib_ls[[10]])
l <- plot_distrib_zeta(x,"Turkish")
dev.off()


```



### MSE of the models

```{r}
MSE()
```































# Alternative distribution: Altman's function

```{r}
get_mle_pars_altman <- function(x) {
      M <- sum(x)
      M_1 <- sum(log(x))
          
      # Loglikelihood
      S <- function(gamma,delta) {
            s <- 0
            for (i in 1:N) {
              s <- s + (i^(-gamma)*exp(-delta*i))
              }
            return(s) 
      }
      
      minus_log_like_alt <- function(gamma,delta) {
            gamma*M_1 + delta*M + log(S(gamma,delta))
      }
      
      mle_alt <- mle(minus_log_like_alt, start = list(gamma = 2, delta = -3),method = "L-BFGS-B")
      
      return(list(attributes(summary(mle_alt))$coef[1],
                  attributes(summary(mle_alt))$coef[2],
                  attributes(summary(mle_alt))$m2logL))
}

x <- unlist(degree_distrib_ls[[1]])
alt_pars <- get_mle_pars_altman(x)
get_AIC(alt_pars[[3]][1],2,N)
```


```{r}
aic_table <- data.frame()

for (n in 1:length(degree_distrib_ls)) {
    x <- unlist(degree_distrib_ls[[n]])
    new_pars <- get_mle_params(x)
    alt_pars <- get_mle_pars_altman(x)
    
    N <- length(x)
    M <- sum(x)
    M_1 <- sum(log(x))
    minus_log_likelihood_zeta2 <- 2*M_1 + N * log((pi^2)/6)
    
    aic_poiss <- get_AIC(new_pars[[2]][1],1,N)
    aic_geom <- get_AIC(new_pars[[2]][2],1,N)
    aic_zeta <- get_AIC(new_pars[[2]][3],1,N)
    aic_trunc_zeta <- get_AIC(new_pars[[2]][4],2,N)
    aic_zeta_2 <- get_AIC(2*minus_log_likelihood_zeta2,0,N)
    aic_alt <- get_AIC(alt_pars[[3]][1],2,N)
    
    aic_ls <- c(aic_poiss,aic_geom,aic_zeta,aic_trunc_zeta,aic_zeta_2,aic_alt)
    aic_table <- rbind(aic_table,aic_ls)
    aic_table$best[n] <- which.min(aic_table[n,1:6])
}

aic_table$lang <- source$language
colnames(aic_table) <- c("Poisson","Geometric","Zeta","Trunc_Zeta","Zeta_2","Altman","best","lang")
aic_table
```





### compare with real distributions
```{r}
degree_sequence = read.table("./samples_from_discrete_distributions/data/sample_of_geometric_with_parameter_0.2.txt",
                             header = FALSE)
```









